---
title: "OneMapSGAPI R Package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{onemapsgapi_vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
library(knitr)
library(kableExtra)
opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

NOT_CRAN <- identical(tolower(Sys.getenv("NOT_CRAN")), "true")
opts_chunk$set(purl = NOT_CRAN)
```

The OneMapSGAPI package provides useful wrappers for the [OneMapSG API](https://docs.onemap.sg/#introduction) client. It allows users to easily query spatial data from the API in a tidy format and provides additional functionalities to allow easy data manipulation. 

```{r, warning = FALSE}
library(onemapsgapi)
```

## Authentication
This function is a wrapper for the [Authentication API endpoint](https://docs.onemap.sg/#authentication-service-post) In order to query data, most API endpoints in the OneMapSG API require a token. First-time users can register themselves using the [OneMapSG registration form](https://developers.onemap.sg/signup/). Subsequently, they can retrieve their tokens using the `get_token()` function with their email and password, for example:

```{r, eval = FALSE}
token <- get_token("user@example.com", "password")
```

```{r, echo = FALSE}
token <- get_token(Sys.getenv("onemap_email"), Sys.getenv("onemap_pw"))
```

The function will also print a message informing users of the token's expiry date and time.

## Themes
These functions are wrappers for the [Themes API](https://docs.onemap.sg/#themes) endpoints. Themes in the OneMap SG API refer to types of locations, such as kindergartens, parks, hawker centres etc.

### Find Themes of Interest
The `search_themes()` function allows users to find details of themes of interest. It returns a tibble of themes matching user's search terms. Alternatively, if no search terms are added, a tibble of all themes available through the API is returned. The variable **THEMENAME** in the output tibble serves as the input for getting theme data.

```{r, eval = NOT_CRAN}
# return all themes related to "hdb" or "parks"
search_themes(token, "hdb", "parks") %>%
  kable() %>% kable_styling() %>%
  scroll_box(height = "300px", width = "100%")
```

```{r, eval = NOT_CRAN}
# return all possible themes
search_themes(token) %>%
  kable() %>% kable_styling() %>%
  scroll_box(height = "300px", width = "100%")
```

```{r, eval = NOT_CRAN}
search_themes(token, more_info = FALSE) %>%
  kable() %>% kable_styling() %>%
  scroll_box(height = "300px")
```

If an error occurs, the function returns `NULL`, along with a warning message.
```{r, eval = NOT_CRAN}
search_themes("my_invalid_token")
```

### Get Theme Data
The function `get_theme()` returns data related to a particular theme, often location coordinates and other information. It returns the output as a tibble, or prints a warning message when an error is encountered. All tibbles will contain the variables: **NAME**, **DESCRIPTION**, **ADDRESSPOSTALCODE**, **ADDRESSSTREETNAME**, **Lat**, **Lng**, **ICON_NAME**, and some provide additional information; for example, query hawker centres gives additional information about the completion date of each hawker centre.

```{r, eval = NOT_CRAN}
# return all hotel data
get_theme(token, "hotels") %>%
  kable() %>% kable_styling() %>%
  scroll_box(height = "300px", width = "100%")
```

```{r, eval = NOT_CRAN}
# return all monuments data within a bounding area
get_theme(token, "monuments", extents = "1.291789,%20103.7796402,1.3290461,%20103.8726032") %>%
  kable() %>% kable_styling() %>%
  scroll_box(height = "300px", width = "100%")
```

```{r, eval = NOT_CRAN}
# returns a list of status tibble and output tibble
get_theme(token, "lighting", return_info = TRUE) %>%
  str()
```

```{r, eval = NOT_CRAN}
# error: output is NULL, warning message shows status code
get_theme("invalid_token", "hotels")

# error: output is NULL, warning message shows error message from request
get_theme(token, "non-existent-theme")

# error: output is query_info, warning message query did not return any records
get_theme(token, "ura_parking_lot", "1.291789,%20103.7796402,1.3290461,%20103.8726032") %>%
  kable() %>% kable_styling() %>%
  scroll_box(width = "100%")
```

## Population Query
These functions are a wrapper for the [Population Query API](https://docs.onemap.sg/#population-query). Population Query API endpoints allow users to pull socio-economic datasets by planning area, which each endpoint representing a dataset (e.g. `getPopulationAgeGroup` provides age group summary statistics by planning area). This package combines querying different Popquery API endpoints into single functions. 

### Specific Query
The function `get_pop_query()` allows users to query a specific dataset for a specific town, year and gender (optional). It is faster for users who only require a specific value. 

The function returns a tibble with 1 row and values for all the corresponding variables returned by the API endpoint.

```{r, eval = NOT_CRAN}
# example: return occupation summary data for Bedok town in year 2010
get_pop_query(token, "getOccupation", "Bedok", "2010") %>%
  kable() %>% kable_styling() %>%
  scroll_box(width = "100%")
```

The `gender` parameter is only valid for the **getEconomicStatus**, **getEthnicGroup**, **getMaritalStatus** and **getPopulationAgeGroup** endpoints. If specified for other endpoints, the parameter will be dropped.

If gender is not specified for endpoints with a gender parameter, records for total, male and female will be returned. The notable exception to this is for the "getEthnicGroup" endpoint, which only returns the total record if gender is not specified. This is because by default, this is the only API endpoint with a gender parameter that does not return gender breakdown by default.

```{r, eval = NOT_CRAN}
# example: endpoints with a gender parameter but gender is not specified
get_pop_query(token, "getEthnicGroup", "Bedok", "2010") %>%
  kable() %>% kable_styling()

get_pop_query(token, "getMaritalStatus", "Bedok", "2010") %>%
  kable() %>% kable_styling()
```

```{r, eval = NOT_CRAN}
# example: endpoints with a gender parameter but gender specified
get_pop_query(token, "getEconomicStatus", "Yishun", "2010", "female")
```

Finally, if an error occurs, the function will return `NULL` and a warning message.

```{r, eval = NOT_CRAN}
# example: returns NULL, warning message shows status code
get_pop_query("invalid_token", "getOccupation", "Bedok", "2010")

# example: returns NULL, warning message shows the error
get_pop_query(token, "getInvalidData", "Bedok", "2010")
get_pop_query(token, "getOccupation", "Bedok", "fakeyear")
get_pop_query(token, "getReligion", "faketown", "2010")
```

### Multiple Queries
The function `get_pop_queries()` allows users to query multiple datasets for multiple towns and years. It is similar to get_pop_query but allows for querying of multiple data types for multiple towns and years. Note that rate-limiting is essential for large queries.

The function returns a tibble with each row representing a town in a particular year for a particular gender, and columns with the variables returned by the API endpoint. If any API call returns no data, the values will be NA but the row will be returned. However, if all data_types do return data for that town and year, no row will be returned for it.

```{r, eval = NOT_CRAN}
# example: returns output with no NA
get_pop_queries(token, c("getOccupation", "getLanguageLiterate"), 
                c("Bedok", "Yishun"), "2010") %>%
  kable() %>% kable_styling() %>%
  scroll_box(width = "100%")

# example: returns output with no NA and gender field
get_pop_queries(token, c("getEconomicStatus", "getEthnicGroup"), 
                "Yishun", "2010", "female") %>%
  kable() %>% kable_styling()
```

If data types requested is a mix of those that accept gender parameters and does that do not, only `gender = "Total"` rows will have all records. The data types that does not accept gender params will be in `gender = Total`. 

```{r, eval = NOT_CRAN}
# example: gender not specified
get_pop_queries(token, c("getEconomicStatus", "getOccupation", "getLanguageLiterate"),
                "Bedok", "2010") %>%
  kable() %>% kable_styling() %>%
  scroll_box(width = "100%")
```

```{r, eval = NOT_CRAN}
# example: gender specified
get_pop_queries(token, c("getEconomicStatus", "getOccupation", "getLanguageLiterate"),
                "Bedok", "2010", gender = "female") %>%
  kable() %>% kable_styling() %>%
  scroll_box(width = "100%")
```

If all data_types do return data for that town and year, no row will be returned for it.
```{r, eval = NOT_CRAN}
# example: no records for 2012
get_pop_queries(token, c("getEconomicStatus", "getOccupation"),
                "Bedok", c("2010", "2012")) %>%
  kable() %>% kable_styling() %>%
  scroll_box(width = "100%")
```


## Route Service
These functions are a wrapper for the [Route Service API](https://docs.onemap.sg/#route). The Route Service API provides users a way to query the route taken from one point to another. It provides information about the total time and distance taken for the route, route instructions and other infomation e.g. elevation, for a variety of routes (public transport, drive, walk, cycle). This package provides three different functions associated with this API, each serving different purposes. 

### All Route Information
The `get_route()` function returns all API output but with standardized column names, which allows for subsequent merging if desired. This is particularly useful as API output variable names may vary depending on parameters (e.g. start point is named differently between `route = drive` and `route = pt`).

It returns the full route data in a tibble format, or a list containing a tibble of results and list of status information if desired.

```{r, eval = NOT_CRAN}
# example: only route data, route = drive
get_route(token, c(1.319728, 103.8421), c(1.319728905, 103.8421581), "drive") %>%
  kable() %>% kable_styling() %>%
  scroll_box(width = "100%")
```

```{r, eval = NOT_CRAN}
# example: only route data, route = pt
get_route(token, c(1.319728, 103.8421), c(1.319728905, 103.8421581), "pt",
          mode = "bus", max_dist = 300) %>%
  kable() %>% kable_styling() %>%
  scroll_box(width = "100%")
```

```{r, eval = NOT_CRAN}
# example: returns list of status list and output tibble
get_route(token, c(1.319728, 103.8421), c(1.319728905, 103.8421581), 
          "drive", status_info = TRUE) %>%
  str()
```

```{r, eval = NOT_CRAN}
# example: error, warning message shows status code
get_route("invalid_token", c(1.319728, 103.8421), c(1.319728905, 103.8421581), "drive")

# example: error, warning message shows error message from request
get_route(token, c(300, 300), c(400, 500), "cycle")
get_route(token, c(1.319728, 103.8421), c(1.319728905, 103.8421581), "fly")
```

### Total time and distance
The function `get_summ_route()` accepts the same parameters as `get_route()`, but only returns total time and total (and start and end points, if requested). Recognising that this API is most valuable for calculating total time travelled (as a improved measure of spatial distance compared to Euclidean distance), this function produces a cleaner output containing only the main variables of interest.

```{r, eval = NOT_CRAN}
# example: drive/walk/cycle
get_summ_route(token, c(1.319728, 103.8421), c(1.319728905, 103.8421581), "drive") %>%
  kable() %>% kable_styling() %>%
  scroll_box(width = "100%")
```

If route = "pt", only the best route is chosen (i.e. n_itineraries = 1).
```{r, eval = NOT_CRAN}
# example: pt
get_summ_route(token, c(1.319728, 103.8421), c(1.319728905, 103.8421581), "pt",
               mode = "bus", max_dist = 300) %>%
  kable() %>% kable_styling() %>%
  scroll_box(width = "100%")
```

If an error occurs, the output will be a tibble of NAs, along with a warning message
```{r, eval = NOT_CRAN}
# example: error, warning message shows status code
get_summ_route("invalid_token", c(1.319728, 103.8421), c(1.319728905, 103.8421581), "drive")

# example: error, warning message shows error message from request
get_summ_route(token, c(300, 300), c(400, 500), "cycle")
get_summ_route(token, c(1.319728, 103.8421), c(1.319728905, 103.8421581), "fly")
```

### Total time and distance matrix
The function `compare_routes()` allows the calculation of total time and distance for a tibble of start and end points. Users input a tibble of start and end points (and potentially other variables) and the function returns a tibble with additional columns, `total_time` and `total_dist`. The function also accepts multiple arguments for `route` and `pt_mode`, allowing users to compare various route options.

```{r, eval = NOT_CRAN}
# example: create sample df with start and end coordinates
sample <- data.frame(start_lat = c(1.3746617, 1.3567797, 1.3361976, 500),
    start_lon = c(103.8366159, 103.9347695, 103.6957732, 501),
    end_lat = c(1.429443081, 1.380298287, 1.337586882, 601),
    end_lon = c(103.835005, 103.7452918, 103.6973215, 600),
    add_info = c("a", "b", "c", "d"))
```

```{r, eval = NOT_CRAN}
# example: multiple routes
compare_routes(token, sample[1:3, ],
    "start_lat", "start_lon", "end_lat", "end_lon",
    routes = c("cycle", "walk")) %>%
  kable() %>% kable_styling() %>%
  scroll_box(width = "100%")
```

```{r, eval = NOT_CRAN}
# example: multiple routes + multiple pt modes 
compare_routes(token, sample[1:3, ],
    "start_lat", "start_lon", "end_lat", "end_lon",
    routes = c("drive", "pt"), pt_mode = c("bus", "transit")) %>%
  kable() %>% kable_styling() %>%
  scroll_box(width = "100%")
```

By default, the data appears in a wide format, but users can specify for the output to be in long format.
```{r, eval = NOT_CRAN}
# example: long format
# no error, long format
compare_routes(token, sample[1:3, ],
    "start_lat", "start_lon", "end_lat", "end_lon",
    routes = c("walk", "pt"), pt_mode = c("bus", "transit"),
    as_wide = FALSE) %>%
  kable() %>% kable_styling()
```

If an error occurs, the output row will be have NAs for the additional variables, along with a warning message.
```{r, eval = NOT_CRAN}
# example: with error
compare_routes(token, sample,
    "start_lat", "start_lon", "end_lat", "end_lon",
    routes = c("cycle", "walk")) %>%
  kable() %>% kable_styling()
```
